---
title: "Reproducible Research - Peer Assessment 1"
author: "Miriam P."
date: "21/01/2018"
output:
  pdf_document: default
  html_document:
    theme: flatly
---

## <a name="top"></a>Overview

This document describes an analysis done for an assignment of the Coursera Reproducible Research Course from Johns Hopkins University. 

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These IoT devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

Click on links here to quickly view tasks completed in this assignment:

1.  [Code for reading in the dataset and/or processing the data](#point1) 
2.  [Histogram of the total number of steps taken each day](#point-2) 
3.  [Mean and median number of steps taken each day](#point-3) 
4.  [Time series plot of the average number of steps taken](#point-4) 
5.  [The 5-minute interval that, on average, contains the maximum number of steps](#point-5) 



The data for this assignment comes from from the course web site:

*   [activityData](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip )

## Details

Before beginning the project, be sure to load the required R libraries and set any environmental variables. Note that setting messages in markdown to false 
suppresses messages from library loading such as version number and dependencies.
Updating to latest versions of these libraries may improve ability to obtain
results fairly similar to the steps outlined here.

```{r, echo=TRUE, results='hide', warning=FALSE, error=FALSE, message=FALSE, warning=FALSE}
# Clear environment of prior calculations. For best results always create/use an R project to run this.
    rm(list=ls())
```

```{r, collapse=FALSE, echo=TRUE, results='markup', error=FALSE, message=FALSE, warning=FALSE}
# load libraries
library(data.table)
library(dplyr)
library(knitr)
library(ggplot2)
library(ggthemes)
library(Hmisc)



```

### <a name="point1"></a>1.    Load and Clean Data


```{r, collapse=FALSE, echo=TRUE, results='markup', error=FALSE, message=FALSE, warning=FALSE}
#### A.  Loading data
# set working environment
    workDataDirStr <- paste(getwd(),"/", sep="")
# Load the files. Using NA function to assure clean file
    read.activity <- function(file) {
        fread(file, na.strings=c("NA", ""))
    } 
    activityData <- read.activity(paste(workDataDirStr, "activity.csv", sep=""))
```

*   Dimensions of activityData (rows,columns): `r dim(activityData)`

Add any steps here to process, scale or transform the data further before analysis:

```{r message=FALSE, warning=FALSE}
#### B.  Cleaning data
# convert date string to date datatype
    activityData$date<-as.Date(activityData$date)
# Since have two libraries with wday function specify lubridate to add Weekday
    activityData$Weekday<-lubridate::wday(activityData$date, label = TRUE, abbr = FALSE)
# convert all steps to whole number
    activityData$steps<-round(activityData$steps, digits=0)
```

 
-----

### <a name="point-2"></a>2. Make a histogram of the total number of steps taken each day

```{r, collapse=FALSE, echo=TRUE, results='markup', error=FALSE, message=FALSE, warning=FALSE}
# Data Exploration - What is mean total number of steps taken per day, per weekday, per interval?

#### A.  Pre-processing
# Query and convert dataset for plotting
    activityData.sums <- activityData %>%
    group_by(date,Weekday) %>%
    summarise(totalSteps = sum(steps, na.rm=TRUE))
# Get max number of steps for largest interval for extending y-axis to fit labels.
    activityData.max <- round(max(activityData.sums$totalSteps, na.rm = TRUE),digits=0)
# Get initial averaging data for histogram plot
    activityData.mean <- round(mean(activityData.sums$totalSteps, na.rm = TRUE),digits=0)
    activityData.median <- round(median(activityData.sums$totalSteps, na.rm = TRUE),digits=0)
```

```{r, error=FALSE, message=FALSE, warning=FALSE}
#### B.  Plotting
# Prepare labels for plot 
    textlab_1 <-  paste("Mean = ", activityData.mean, sep="")
    textlab_2 <-  paste("Median = ", activityData.median, sep="")
# Define and call plot
    plot_2 <- ggplot(data=activityData.sums, aes(activityData.sums$totalSteps)) + 
        geom_histogram(
                 col="black", 
                 aes(fill=..count..),
                 alpha = .8,
                 binwidth = 2000) +
        labs(title = "Frequency Distribution Histogram of Step Activity",
        y="% Frequency",
        x="Total Steps per Day") +
        theme_economist() +
        geom_vline(xintercept = activityData.mean, colour="blue", linetype = "solid") +
        geom_vline(xintercept = activityData.median, colour="red", linetype = "longdash") +
        annotate("text", label = textlab_1, x = activityData.mean, y = 15, size = 5, colour = "blue") +
        annotate("text", label = textlab_2, x = activityData.median, y = 13, size = 5, colour = "red") +
        theme(legend.position = "none") + 
        scale_fill_gradient("Count", low = "white", high = "gray")  +
        scale_colour_economist()
    plot1

```
 

 
-----

### <a name="point-3"></a>3. Calculate and report the mean/median total steps taken per day

```{r}
#### A.  Calculate and Display initial daily averaging data
    dailyAvgSteps <- activityData.sums %>%
    group_by(Weekday,totalSteps) %>%
    filter(!is.na(totalSteps))

    dailyAvgSteps           #Preveview of daily average ateps

#### B. Describe norms          
    dailyAvgSteps.mean   <- round(mean(dailyAvgSteps$totalSteps, na.rm = TRUE), digits=0)
    dailyAvgSteps.mean      # Mean number of steps
    
    dailyAvgSteps.median <- round(median(dailyAvgSteps$totalSteps, na.rm = TRUE), digits=0)
    dailyAvgSteps.median    # Median number of steps
   
```

Answer: The plot shows that the mean is lower than the median indicating a
potential issue with missing data that could impact calculations on this set.

 [back to top](#top)
 
-----

### <a name="point-4"></a>4. Display a time series plot

```{r, collapse=FALSE, echo=TRUE, results='markup', error=FALSE, message=FALSE, warning=FALSE}
#### A.  What is the average daily activity pattern?
    activityData.intervals <- activityData %>%
    group_by(interval) %>%
    filter(!is.na(steps)) %>%
    summarise(meanSteps = mean(steps, na.rm=TRUE))

    activityData.intervals   # Preview of average steps per tnterval
```

```{r}
#### B.  Prepare a histogram of frequency counts/percentages.
    plot_4 <- ggplot(data=activityData.intervals, aes(x=interval, y=meanSteps)) +
        geom_line(colour="gray", linetype="solid", size=1.5) + 
        labs(title = "Interval Step Activity Drilldown",
             y="Average Steps per Interval",
             x="5-minute interval IDnum")+
        theme_economist() +
        scale_colour_economist()
    plot2
```
 
 [back to top](#top) 
 
-----

###  <a name="point-5"></a>5. Determine average daily maximum steps in a 5-minute interval
```{r}
#### A. Find the largest item in series and call related descriptive informat
    mostInt <- which.max(activityData.intervals$meanSteps)
    intervalMax <- activityData.intervals[mostInt,]
    maxIntID <- intervalMax$interval
    maxSteps <- round(intervalMax$meanSteps, digits=0)
```
Answer: The most number of steps occurred in the morning

*   The number of Most Steps: `r maxSteps`
*   Occurred at Interval ID (time of day): `r maxIntID`

 [back to top](#top) 
 
-----
